Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-18 paper/high-assurance-contracts.tex
Update text in paper to match contract changes.
Reformat filter to be consistent with other contracts
Expunging dialogue diagram

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \includegraphics[scale=0.3]{dialogue.png}
    \end{tabular}
  \end{center}
  \caption{BriefCASE dialogue for filter transformation.}
  \label{fig:dialogue}
\end{figure}

@@

    guarantee Filter_Output
    "Filter output is well-formed" :
    if event(Input) and policy then 
        event(Output) and Output = Input
    else not event(Output);
  \end{lstlisting}
\end{lrbox}

@@

    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
                         (    (isPending => latency < MAX_LATENCY)   
                          and (rsp => (req or pre(isPending))));
    -- stop user definitions

@@

As noted previously, the system with the untrusted AI does not pass AGREE verification.
Transformations in BriefCASE are used to cyber-harden the system against malicious AI behavior by inserting a high-assurance filter and monitor between the AI and the WM as shown in \figref{fig:hardened}.
High-assurance means the behaviors of the components are verified.
Here the filter protects the WM from malformed data and the monitor protects it from spontaneous or delayed responses.
These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear and then choosing the appropriate transformation.

The system designer provides configuration parameters for the transformation in a dialogue box.
The dialogue box for adding a filter is shown in \figref{fig:dialogue}.
All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
A filter policy defines well-formed data while a monitor policy defines an invariant over inputs and outputs.
These policies can be stated directly in the wizard, or they can be left blank and added later to the AGREE contract generated by the transformation.

BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
It also creates a default AGREE specification that defines everything about the component behavior but the policy itself unless it is provided in the dialogue.
The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
The single guarantee completely defines the filter output under any input according to the truth value of the policy.
While such a strong guarantee is not required for AGREE analysis, it is required for automated synthesis.

The AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerably more complex than the filter.
BriefCASE generates everything but the statements in \linesref{line:mon-assume}{line:mon-policy}.
As with the filter, the guarantees completely define the output, as required for synthesis, in terms of the policy and the alert.
The \texttt{alert} variable defined on \lineref{line:mon-alert} depends not just on the policy but also on the indicated behavior from the BriefCASE dialogue box. 
Marking \texttt{is\_latched} as true makes the alert persistent, meaning that once the alert is raised, it is always raised; otherwise, it is the complement of the policy value in the current step.

>>>>

%% \begin{figure}
%%   \begin{center}
%%     \begin{tabular}{c}
%%     \includegraphics[scale=0.3]{dialogue.png}
%%     \end{tabular}
%%   \end{center}
%%   \caption{BriefCASE dialogue for filter transformation.}
%%   \label{fig:dialogue}
%% \end{figure}

@@

    guarantee "Filter output is well-formed" :
      if event(Input) and policy then
        event(Output) and Output = Input
      else
        not event(Output);
  \end{lstlisting}
\end{lrbox}

@@

    eq policy : bool = (rsp => req) ->
                         (    (isPending => latency < MAX_LATENCY)
                          and (rsp => (req or pre(isPending))));
    *\label{line:mon-policy}*
    -- stop user definitions

@@

As noted previously, the system with the untrusted AI does not pass
AGREE verification.  Transformations in BriefCASE cyber-harden the
system against malicious AI behavior by inserting a high-assurance
filter and monitor between the AI and the WM as shown in
\figref{fig:hardened}.  (In our context \emph{high-assurance} means
that the behaviors of the components are verified to meet their
contracts.)  In the example, the inserted filter protects the WM from
malformed data and the monitor protects the WM from (possibly malicious)
spontaneous or delayed responses.  These high-assurance components are
added, one at a time, by
\begin{itemize}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{itemize}

%% The system designer provides configuration parameters for the
%% transformation in a dialogue box.  The dialogue box for adding a
%% filter is shown in \figref{fig:dialogue}.  All high-assurance
%% components rely on a \emph{policy} to define behavior as seen in the
%% last field of the dialogue box.  A filter policy defines well-formed
%% data while a monitor policy defines an invariant over inputs and
%% outputs.  These policies can be stated directly in the wizard, or they
%% can be left blank and added later to the AGREE contract generated by
%% the transformation.

BriefCASE creates all the needed AADL for the new high-assurance
component and its connections in the system implementation as part of
the transformation.  It also creates a default AGREE specification
that defines everything about the component behavior but the policy
itself, unless that is provided in the dialogue.  The resulting AGREE
specification from the transformation for the filter is shown in
\figref{fig:filter}.  The single guarantee completely defines the
filter output under any input according to the truth value of the
policy.  Such a strong guarantee is needed for synthesis (a weaker
version suffices for AGREE analysis).

The AGREE specification for the monitor is shown in
\figref{fig:monitor}. As can be seen, it is considerably more complex
than the filter specification.  BriefCASE generates everything but the
statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
the filter, the guarantees completely define the output, as required
for synthesis, in terms of the policy and the alert.  The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  Setting \texttt{is\_latched} as true
makes the alert persistent, meaning that once the alert is raised, it
is always raised; otherwise, it is the complement of the policy value
in the current step.
